---
title: "Descriptive Analysis"
author: "jvicentem"
date: "26 July 2017"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
source('utils.R')

#devtools::install_github("ujjwalkarn/xda")

library(knitr)

library(ggjoy)

#devtools::install_github("vsimko/corrplot")
library(corrplot)
```

```{r loading-data, message=FALSE, warning=FALSE}
df <- load_data()

df <- introduce_nas(df, 22.5, 'pH')


```

# Data dictionary

https://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality.names

   Input variables (based on physicochemical tests):
   1 - fixed acidity
   2 - volatile acidity
   3 - citric acid
   4 - residual sugar
   5 - chlorides
   6 - free sulfur dioxide
   7 - total sulfur dioxide
   8 - density
   9 - pH
   10 - sulphates
   11 - alcohol
   12 - wine_colour
   Output variable (based on sensory data): 
   13 - quality (score between 0 and 10)
   
4. Relevant Information:

   The two datasets are related to red and white variants of the Portuguese "Vinho Verde" wine.
   For more details, consult: http://www.vinhoverde.pt/en/ or the reference [Cortez et al., 2009].
   Due to privacy and logistic issues, only physicochemical (inputs) and sensory (the output) variables 
   are available (e.g. there is no data about grape types, wine brand, wine selling price, etc.).

   These datasets can be viewed as classification or regression tasks.
   The classes are ordered and not balanced (e.g. there are munch more normal wines than
   excellent or poor ones). Outlier detection algorithms could be used to detect the few excellent
   or poor wines. Also, we are not sure if all input variables are relevant. So
   it could be interesting to test feature selection methods. 

5. Number of Instances: red wine - 1599; white wine - 4898. 

6. Number of Attributes: 11 + output attribute
  
   Note: several of the attributes may be correlated, thus it makes sense to apply some sort of
   feature selection.
   
   
```{r summaries}
xda::numSummary(df)
xda::charSummary(df)
```

```{r plots}
plot_histograms(df)

plot_boxplots(df)

plot_joyplots(df)

other_plots(df)
```

```{r plots2}
boxplots_quality_colour_wine(df, 'red')

boxplots_quality_colour_wine(df, 'white')
```

```{r cors}
corrs <- cor(drop_na(df) %>% select(-wine_colour))

corrs

corrplot.mixed(corrs, upper="ellipse", lower="number")
```



